Application of a Discontinuous Petrov-Galerkin method to compressible flow problems

### 2012-09-07 docs/Jesse/Vienna/pres.tex

\title[A DPG method for Compressible Flow]{A Discontinuous Petrov-Galerkin method for compressible flow problems}
\date{August 15, 2012}
\author[J. Chan, T. Ellis, L. Demkowicz, R. Moser]{Jesse Chan, Truman Ellis \\ Supervisors: Leszek Demkowicz, Robert Moser}

@@

\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers - resolution needed for drag/load
\item{} Turbulence (non-localized)
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}
\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the standard Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Streamline-upwind Petrov-Galerkin (SUPG)}
SUPG solves $b_{\rm SUPG}(u,v) = l_{\rm SUPG}(v)$, where
\begin{align*}
b_{\rm SUPG}(u,v) &= b(u,v)+ \sum_{K} \int_{K} \tau \LRp{L_{\rm adv} v} L u \\
l_{\rm SUPG}(v) &= \ell(v) + \sum_{K} \int_{K} \tau \LRp{L_{\rm adv}v} f.
\end{align*}
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.44\textwidth}
\vspace{-.5cm}
\begin{itemize}
\item{} $Lu = \div\left(\beta u\right) - \epsilon \triangle u$, $L_{\rm adv}u = \div \left(\beta u\right)$, and $\tau$ is a parameter. 
%\item{} Recovers ``exact'' artificial diffusion for $f=0$.
\item{} Effective for $f\neq 0$. %``exact'' artificial diffusion.
\item{} \emph{Residual-based} stabilization. 
\end{itemize}
\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.38]{figs/SUPG.png}
%\caption{Exact, Bubnov-Galerkin, and SUPG solutions.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be interpreted as a \emph{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\bibentry{SUPG}}
}

\section{The DPG Method}
\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the dual functional over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the isometric Riesz map st.\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our functional $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
\end{frame}

\frame{
\frametitle{Summary: select test functions to minimize residuals}
For $\delta u \in U_h$, define the \textcolor{red}{optimal test function} $v_{\delta u}$. 
\begin{equation*}
v_{\delta u} \coloneqq R_V^{-1}B\delta u.
\end{equation*} 
Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
\[
b\LRp{u_h,v_{\delta u}} = \ell\LRp{v_{\delta u}}, \quad \forall \delta u \in U_h.
\]
}

@@

\frame{
\frametitle{Properties of DPG}
DPG provides a symmetric positive-definite stiffness matrix. Let $\{\phi_j\}_{j=1}^m$ be a basis for $U_h$, and $\{v_i\}_{i=1}^n$ a basis for $V_h$, such that $n>m$. Then, for 
\begin{align*}
B_{ji} &= b(\phi_j,v_i),\\
l_i &= \ell(v_i),
\end{align*}
DPG solves the discrete system for degrees of freedom $u$
\begin{align*}
\LRp{B^T R_V^{-1}B} u  &= \LRp{B^T R_V^{-1}} l,
\end{align*}
where, under a localizable norm and discontinuous test functions, $R_V^{-1}$ is block diagonal.
}

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(l-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\end{itemize}
\footnotetext{\bibentry{DPG3pub}}
}

%\frame{
%\frametitle{Duality of norms}
%Duality of norms: for any trial norm $\nor{u}_U$,
%\[
%\nor{u}_{U} = \sup_{v \in V}\frac{b\LRp{u,v}}{\nor{v}_{V,U}}, \quad \nor{v}_{V,U} = \sup_{u \in U}\frac%{b\LRp{u,v}}{\nor{u}_{U}}.
%\]
%Likewise, for any test norm $\nor{v}_V$, 
%\[
%\nor{v}_{V} = \sup_{u \in U}\frac{b\LRp{u,v}}{\nor{U}_{U,V}}, \quad \nor{u}_{U,V} = \sup_{v \in V}\frac%{b\LRp{u,v}}{\nor{v}_{V}}.
%\]
%
%}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\footnote{\bibentry{analysisDPG}}
}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|^2
\] 
generates a test norm \emph{equivalent} to 
\[
\nor{v}_{V_{\rm opt}} \coloneqq \|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute the jump terms for an $L^2$ term, giving us the \emph{graph norm}\footnote{\bibentry{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}.
\]
}

\frame{
\frametitle{Goal: robust higher order adaptive methods}
\begin{columns}
\begin{column}{.45\textwidth}
Our aim is to design an adaptive method for compressible laminar flows that is 
\begin{itemize}
\item robust in Reynolds number
\item stable for arbitrary elements
\end{itemize}

This proposal will outline
\begin{enumerate}
\item a robust DPG method for convection-diffusion,
\item extension to nonlinear problems and systems,
\item proposed work.
\end{enumerate}

\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.23]{figs/Re1000p2/u2_full.png}}
\subfigure{\includegraphics[scale=.37]{figs/Re1000p2/mesh6.png}}
\caption{$u_2$ and $p=2$ mesh for $\Reyn = 1000$.}
\end{figure}
\end{column}
\end{columns}
}

@@

\frame{
\frametitle{Ultra-weak formulation for convection-diffusion}
In first order form, the convection-diffusion equation is
\[
A \LRp{u,\sigma} \coloneqq 
\LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\]
The variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
where $\widehat{f}_n \coloneqq \beta_n u - \sigma_n$ and $\LRa{\widehat{f}_n,\jump{v} }_{\Gh}$ is defined 
\[
\LRa{\widehat{f}_n,\jump{v} }_{\Gh}\coloneqq \sum_K\int_{\partial K}{\rm sgn}(\vec{n}) \, \widehat{f}_n v.
\]
}

\frame{
\frametitle{Graph norm under convection-diffusion}
For convection-diffusion, the graph test norm is defined elementwise
\[
\|\left(v,\tau\right)\|_{V(K)}^2 = \| \div \tau - \beta \cdot \grad v
\|_{L^2(K)}^2 + \| \epsilon^{-1} \tau + \grad v \|_{L^2(K)}^2 +
\|v\|_{L^2(K)}^2.
\]
Problem with this test norm: approximability of test functions.
\vspace{-.3cm}
\begin{figure}[!h]
\centering
\includegraphics[scale=.175]{figs/opt.png}
\caption{$v$ and $\tau$ components of the 1D optimal test functions for flux $\widehat{f}_n$ on the \emph{right-hand} side of a unit element for $\epsilon = 0.01$. }
\label{fig:optTestBoundary}
\end{figure}
}

\frame{
\frametitle{Determining an alternative test norm}
Recall the convection-diffusion bilinear form
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh}&\\
 - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
We recover $\nor{u}^2_{L^2(\Omega)}$ by choosing conforming $\LRp{v,\tau}$ satisfying the \emph{adjoint equations}
\begin{align*}
\div\tau - \beta \cdot \grad v   &= u\\
\frac{1}{\epsilon}\tau + \grad v &= 0
\end{align*}
with boundary conditions s.t.\ $\LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gamma}$ and $\LRa{ \widehat{f}_n,\jump{v} }_{\Gamma}$ vanish.
}

\frame{
\frametitle{A robust bound}
``Necessary'' conditions for robustness --- let $\boldsymbol U = \LRp{u,\sigma,\widehat{u},\widehat{f}_n}$. Then, by choosing specific $\LRp{v,\tau}$ satisfying the adjoint equations,
\[
\nor{u}^2_{L^2(\Omega)} = b\LRp{{\boldsymbol U},\LRp{v,\tau}} = \frac{b\LRp{{\boldsymbol U},\LRp{v,\tau}}}{\nor{\LRp{v,\tau}}_V} \nor{\LRp{v,\tau}}_V \leq \nor{\boldsymbol U}_E \nor{\LRp{v,\tau}}_V
\]
Let $\lesssim$ denote a robust bound - \textcolor{red}{if $ \nor{\LRp{v,\tau}}_V \lesssim \|u\|_{L^2(\Omega)}$}, then we have that
\[
\nor{u}_{L^2(\Omega)} \lesssim \nor{\boldsymbol U}_E
\]
\textbf{Main idea: the \textcolor{red}{test norm} should measure \textcolor{red}{adjoint solutions} robustly.}
}

\frame{
\frametitle{Choice of inflow boundary condition}

\begin{columns}
\begin{column}{.49\textwidth}
We impose the standard outflow wall boundary condition on $u$. For inflow boundary condition: 
\begin{itemize}
\item{} The standard choice of inflow boundary condition: $u = u_0$.
\item{} We impose the non-standard inflow condition: $\widehat{f}_n \coloneqq \beta_n u - \sigma_n \approx \beta_n u_0$ on $\Gamma_{\rm in}$.
\end{itemize}
\end{column}

\begin{column}{.49\textwidth}
\begin{figure}
\includegraphics[scale=.23]{figs/primal.png}
\caption{Non-standard inflow.}
\end{figure}
\end{column}
\end{columns}
When $\sigma_n \approx 0$ near the inflow, condition on $\widehat{f}_n$ approximates condition on $u$.

}

\frame{
\frametitle{Dirichlet inflow boundary condition}
Standard choice of boundary condition: $u = u_0$ on inflow boundary $\Gamma_{\rm in}$, induces boundary layers in adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(\epsilon^{-1})$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primalDir.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjointDir.png}}
\caption{For the standard Dirichlet inflow condition, the solution to the adjoint problem can develop strong boundary layers at the outflow of the adjoint problem. }
\end{figure}
}

\frame{
\frametitle{New inflow boundary condition on $\widehat{f}_n$}
Non-standard choice of boundary condition: $\widehat{f}_n = \beta_n u_0$ on $\Gamma_{\rm in}$, induces smoother adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(1)$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primal.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjoint.png}}
\caption{Under the new inflow condition, the wall-stop boundary condition is relaxed to a zero-stress condition at the outflow boundary of the adjoint problem.}
\end{figure}
}

@@

%\frame{
%\frametitle{``Building blocks'' norm}
%Test norm quantities are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  
%\[
%\|\left(v,\tau\right)\|_{V,K}^2 = \|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \frac{1}{\epsilon}\|\tau\|^2.
%\]
%Problem: boundary layers in optimal test functions:
%\begin{figure}[!h]
%\centering
%\subfigure{\includegraphics[scale=.15]{figs/mesh1.png}}
%\subfigure{\includegraphics[scale=.15]{figs/sol1.png}}
%\caption{The $v$ component of the optimal test function corresponding to flux $\widehat{u} = x(1-x)$ on the bottom side of a unit element for $\epsilon = 0.01$.}
%\label{fig:boundaryTest}
%\end{figure}
%}
\frame{
\frametitle{Adjoint estimates, or ``norm building blocks''}
For solutions $(v,\tau)$ of the adjoint equations, the following quantities are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  
\begin{align*}
&\|v\|, \sqrt{\epsilon} \|\grad v\|, \|\beta \cdot \grad v\| \\
&\| \div \tau\|, \frac{1}{\sqrt{\epsilon}}\|\tau\|.
\end{align*}
We will construct a test norm through a combination of the above quantities, such that
\begin{itemize}
\item $v$ and $\tau$ decoupled (no systems).
\item Coefficients are of equal order after transforming to the unit element (no boundary layers).
\end{itemize}
}
\frame{
\frametitle{Mesh-scaled test norms}
Our test norm, as defined over a single element $K$, is now
\begin{align*}
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 +&\\
\| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2&.
\end{align*}
which induces the proven \emph{robust} bound\footnote{\bibentry{DPGrobustness2}}
\[
\nor{u}_{L^2(\Omega)}+ \nor{\sigma}_{L^2(\Omega)} + \epsilon\nor{\widehat{u}} + \sqrt{\epsilon} \nor{\widehat{f}_n} \lesssim \nor{\LRp{u,\sigma,\widehat{u},\widehat{f}_n}}_E.
\]
}

\frame{
\frametitle{Erikkson-Johnson model problem}

On domain $\Omega = [0,1]^2$, with $\beta = (1,0)^T$, $ f = 0$ and boundary conditions
\begin{align*}
\widehat{\beta_nu - \sigma_n} &= \widehat{f}_n = u_0, \quad \beta_n \leq 0\\
\widehat{u} &= 0, \quad \beta_n > 0
\end{align*}

>>>>

\title[A DPG method for Compressible Flow]{Application of a Discontinuous Petrov-Galerkin method to compressible flow problems}
\date{Sept 12, 2012}
\author[J. Chan, L. Demkowicz, R. Moser]{Jesse Chan\\ Supervisors: Leszek Demkowicz, Robert Moser}

@@

\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.55\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers - resolution needed for drag/load
\end{itemize}
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\item{} Stabilized methods
\begin{itemize}
\item{} Stabilization parameters
\item{} Higher-order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.4\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale = .6]{figs/compFlowPics/airfoil_mesh.png}}
\subfigure{\includegraphics[scale=.3]{figs/SUPG.png}}
\end{figure}
\end{column}
\end{columns}
}

\section{The DPG Method}
\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v) \Longleftrightarrow Bu = \ell, \quad 
\left.\begin{array}{cl}
\langle Bu,v\rangle_V &\coloneqq b(u,v)\\
\langle \ell,v\rangle_V &\coloneqq \ell(v).
\end{array}\right\}
\]
We seek to minimize the \textcolor{red}{dual residual} over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 = \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\]
where $R_V: V \to V'$ is the isometric Riesz map
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Conditions for minimization of the convex functional give
\[
b(u_h,v_{\delta u}) = \ell(v_{\delta u}), \quad \forall \delta u \in U_h, \quad v_{\delta u} \coloneqq R_V^{-1}B\delta u.
\]
}

@@

\frame{
\frametitle{Properties of DPG}
%Properties of DPG include\footnotemark
\begin{itemize}
\item Stiffness matrices are \textcolor{red}{symmetric positive-definite}. For trial/test bases $\{\phi_j\}_{j=1}^m$ and $\{v_i\}_{i=1}^n$, with $B_{ji} = b(\phi_j,v_i)$ and $l_i = \ell(v_i)$. DPG solves 
\begin{align*}
\LRp{B^T R_V^{-1}B} u  &= \LRp{B^T R_V^{-1}} l,
\end{align*}
For localizable norms and discontinuous testing, $R_V$ is block diagonal.
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the \textcolor{red}{error representation function} $e$ defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(l-Bu_h)}_V = \nor{e}_V
\]
\end{itemize}
%\footnotetext{\bibentry{DPG3pub}}
}

@@

\frame{
\frametitle{Ultra-weak formulation for convection-diffusion}
\begin{columns}
\begin{column}{.65\textwidth}
The first order convection-diffusion system:
\[
A \LRp{u,\sigma} \coloneqq 
\LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\]
\end{column}

\begin{column}{.35\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[scale = .15]{figs/mesh_skel.pdf}
\end{figure}
\end{column}

\end{columns}

The variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
where $\widehat{f}_n \coloneqq \beta_n u - \sigma_n$ and $\LRa{\widehat{f}_n,\jump{v} }_{\Gh}$ is defined 
\[
\LRa{\widehat{f}_n,\jump{v} }_{\Gh}\coloneqq \sum_K\int_{\partial K}{\rm sgn}(\vec{n}) \, \widehat{f}_n v.
\]
}

\frame{
\frametitle{Graph norm under convection-diffusion}
The graph norm\footnote{\bibentry{Bui-ThanhDemkowiczGhattas11b}} for convection-diffusion gives exceptional stability.  %, this norm is defined elementwise
\begin{align*}
\|\left(v,\tau\right)\|_{V(K)}^2 =& \| \div \tau - \beta \cdot \grad v\|_{L^2(K)}^2 + 
\| \epsilon^{-1} \tau + \grad v \|_{L^2(K)}^2 +
\|v\|_{L^2(K)}^2.
\end{align*}
Problem with this test norm: approximability of test functions.
\begin{figure}[!h]
\centering
\includegraphics[scale=.16]{figs/opt.png}
\caption{Components of optimal test functions for flux $\widehat{f}_n$ on the \emph{right-hand} side of a unit element for $\epsilon = 0.01$. }
\label{fig:optTestBoundary}
\end{figure}
}

\frame{
\frametitle{Determining an alternative test norm}
%Recall the convection-diffusion bilinear form.  %Let $\boldsymbol U = \LRp{u,\sigma,\widehat{u},\widehat{f}_n}$. 
\vspace{-.85cm}
\begin{align*}%b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),\left( v, \tau \right)\right) 
b(\boldsymbol U,\boldsymbol V) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh} + \text{boundary terms}
% - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}
\end{align*}
Recover $\nor{u}^2_{L^2(\Omega)}$ with conforming $\LRp{v,\tau}$ satisfying the \emph{adjoint equations}
\begin{align*}
\left.
\begin{array}{cl}
\div\tau - \beta \cdot \grad v   &= u\\
\frac{1}{\epsilon}\tau + \grad v &= 0
\end{array}
\right. ,
\quad
\LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gamma}, \LRa{ \widehat{f}_n,\jump{v} }_{\Gamma} = 0
\end{align*}
``Necessary'' conditions for robustness --- 
\[
\nor{u}^2_{L^2(\Omega)} = b\LRp{{\boldsymbol U},\LRp{v,\tau}} = \frac{b\LRp{{\boldsymbol U},\LRp{v,\tau}}}{\nor{\LRp{v,\tau}}_V} \nor{\LRp{v,\tau}}_V \leq \nor{\boldsymbol U}_E \nor{\LRp{v,\tau}}_V
\]
Let $\lesssim$ denote a robust bound - \textcolor{red}{if $ \nor{\LRp{v,\tau}}_V \lesssim \|u\|_{L^2(\Omega)}$}, then we have that
\[
\nor{u}_{L^2(\Omega)} \lesssim \nor{\boldsymbol U}_E
\]
\textbf{Main idea: the \textcolor{red}{test norm} should measure \textcolor{red}{adjoint solutions} robustly.}
}

%\frame{
%\frametitle{Choice of inflow boundary condition}
%
%\begin{columns}
%\begin{column}{.49\textwidth}
%We impose the standard outflow wall boundary condition on $u$. For inflow boundary condition: 
%\begin{itemize}
%\item{} The standard choice of inflow boundary condition: $u = u_0$.
%\item{} We impose the non-standard inflow condition: $\widehat{f}_n \coloneqq \beta_n u - \sigma_n \approx \beta_n u_0$ on $\Gamma_{\rm in}$.
%\end{itemize}
%\end{column}
%
%\begin{column}{.49\textwidth}
%\begin{figure}
%\includegraphics[scale=.23]{figs/primal.png}
%\caption{Non-standard inflow.}
%\end{figure}
%\end{column}
%\end{columns}
%When $\sigma_n \approx 0$ near the inflow, condition on $\widehat{f}_n$ approximates condition on $u$. 
%}

\frame{
\frametitle{Dirichlet inflow condition: issues as $\epsilon \rightarrow 0$}
Standard choice of boundary condition: $u = u_0$ on inflow boundary $\Gamma_{\rm in}$, induces boundary layers in adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(\epsilon^{-1})$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primalDir.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjointDir.png}}
\caption{For the standard Dirichlet inflow condition, the solution to the adjoint problem can develop strong boundary layers at the outflow of the adjoint problem. }
\end{figure}
}

\frame{
\frametitle{Solution: New inflow boundary condition on $\widehat{f}_n$}
Non-standard choice of boundary condition: $\widehat{f}_n = \beta_n u_0$ on $\Gamma_{\rm in}$, induces smoother adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(1)$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primal.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjoint.png}}
\caption{Under the new inflow condition, the wall-stop boundary condition is relaxed to a zero-stress condition at the outflow boundary of the adjoint problem.}
\end{figure}
}

@@

\frame{
\frametitle{Mesh-scaled test norms}
For solutions $(v,\tau)$ of the adjoint equations, we derive quantities that are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  Our test norm, as defined over a single element $K$, is now
\begin{align*}
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 +&\\
\| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2&.
\end{align*}
which induces the proven \emph{robust} bound\footnote{\bibentry{DPGrobustness2}}
\[
\nor{u}_{L^2(\Omega)}+ \nor{\sigma}_{L^2(\Omega)} + \epsilon\nor{\widehat{u}} + \sqrt{\epsilon} \nor{\widehat{f}_n} \lesssim \nor{\LRp{u,\sigma,\widehat{u},\widehat{f}_n}}_E.
\]
}

\frame{
\frametitle{Numerical verification: Erikkson-Johnson problem}

%On domain $\Omega = [0,1]^2$, with $\beta = (1,0)^T$, $ f = 0$ and boundary conditions
%\begin{align*}
%\widehat{\beta_nu - \sigma_n} &= \widehat{f}_n = u_0, \quad \beta_n \leq 0\\
%\widehat{u} &= 0, \quad \beta_n > 0
%\end{align*}
Experiments done using Camellia\footnote{\bibentry{Camellia}} and a boundary layer solution.
%\begin{figure}
%\centering
%\subfigure{
%\includegraphics[scale=.23]{figs/wallBC_exact_u.png}
%}
%\subfigure{
%\includegraphics[scale=.23]{figs/wallBC_exact_sigx.png}
%}
%\subfigure{
%\includegraphics[scale=.23]{figs/wallBC_exact_sigy.png}
%}
%\caption{Exact solution for $u$, $\sigma_x$, and $\sigma_y$ for $\epsilon = .01$, $C_1 = 1$, $C_n=0$, $n\neq 1$}
%\end{figure}
\begin{figure}
\centering
\subfigure{
\includegraphics[scale=.33]{figs/errorrates_wallBC.png}
}
\subfigure{
\includegraphics[scale=.33]{figs/L2energyratio_wallBC.png}
}
%\subfigure{
%\includegraphics[scale=.22]{figs/u_pointdiff_wallBC.png}
%}
\caption{$L^2$/energy errors and their ratio for $\epsilon = .01, .001, .0001$.}
\end{figure}
}
